
In this question we have explored below five asset classes:

- US domestic equities (SPY: the S&P 500 stock index)
- US Treasury bonds (TLT)
- Investment-grade corporate bonds (LQD)
- Emerging-market equities (EEM)
- Real estate (VNQ)

We have taken our data from 2007-01-01. After exploring the data we created three portfolios for these assets assuming we have notional $100,000 to invest in one of these portfolios.

1)
Firstly, we explored the properties of five assets we are considering.

Installing relevant libraries. _Quantmod_ helps us to download several years of daily data on these ETFs.
```{r include = FALSE}
library(mosaic)
library(quantmod)
library(foreach)
```


Importing the stocks we are considering
```{r include = FALSE}
mystocks = c("SPY", "TLT", "LQD","EEM","VNQ")
myprices = getSymbols(mystocks, from = "2007-01-01")
```


Adjusting for splits and dividends
```{r}
SPYa = adjustOHLC(SPY)
TLTa = adjustOHLC(TLT)
LQDa = adjustOHLC(LQD)
EEMa=adjustOHLC(EEM)
VNQa=adjustOHLC(VNQ)
```


Looking at close-to-close changes
```{r}
par(bg = "gray")
plot(ClCl(TLTa),main = "US Treasury Daily Returns",ylab = "Return",col = "navy")
plot(ClCl(LQDa),main = "Investment Grade Corporate Bonds Daily Returns",ylab = "Return",col = "navy")
plot(ClCl(SPYa),main = "S&P 500 Daily Returns",ylab = "Return",col = "navy")
plot(ClCl(EEMa),main = "Emerging Market ETF Daily Returns",ylab = "Return",col = "navy")
plot(ClCl(VNQa),main = "REIT Daily Returns",ylab = "Return",col = "navy")
``` 

Combining close to close changes in a single matrix
```{r}
all_returns = cbind(ClCl(SPYa),ClCl(TLTa),ClCl(LQDa),ClCl(EEMa),ClCl(VNQa))
all_returns = as.matrix(na.omit(all_returns))
```

We calculated the sharpe ratio of individual stocks.

```{r}
summary_stocks = cbind(t(t(colMeans(all_returns))),t(t(apply(all_returns,2,sd))))
summary_stocks = cbind(summary_stocks,summary_stocks[,1]/summary_stocks[,2])
colnames(summary_stocks) = c("Average Returns","ST.Dev of Returns","Sharpe Ratio")
summary_stocks
```

These returns can be viewed as draws from the joint distribution
```{r}
pairs(all_returns)
```


The correlation matrix between stocks
```{r}
cor(all_returns)
```

2)
Now we are going to create our first portfolio with even split. We will assign the weight of each asset as 20%. 

```{r}
initial_wealth = 100000
set.rseed(1)    #Setting seed
sim1 = foreach(i=1:5000, .combine='rbind') %do% {
  total_wealth = initial_wealth
  weights = c(0.2, 0.2, 0.2, 0.2, 0.2)
  holdings = weights * total_wealth
  n_days = 20
  wealthtracker = rep(0, n_days)
  
  for(today in 1:n_days) {
    return.today = resample(all_returns, 1, orig.ids=FALSE)
    holdings = holdings + holdings*return.today
    total_wealth = sum(holdings)
    wealthtracker[today] = total_wealth
    #rebalance portfolio
    holdings = weights * total_wealth
  }
  wealthtracker
}

head(sim1)

```

Profit/loss of our portfolio:

```{r}
par(bg = "gray")
mean(sim1[,n_days])
hist(sim1[,n_days]- initial_wealth, breaks=30,main = "Histogram of Portfolio Returns",xlab = "End of Period Return",ylab = "Frequency",col = "blue")
abline(v = mean(sim1[,n_days]) - initial_wealth,col = "red",lw = 3)
```

Calculating 5% value at risk for this portfolio

```{r}
quantile(sim1[,n_days], 0.05) - initial_wealth
```

Hence for this portfolio the average return is $100946.6 and value at risk is $6297.993  


3) Now we are going to create a safer portfolio which reduces our risk. To create this portfolio, we first calculated the average returns and risk involved with each asset had we invested all our wealth into that portfolio. Based on the results we tried different combinations and came up with weights which increased the return of portfolio and minimized the risk involved. Below are the weights:
TLT - 20%
LQD - 30%
SPY - 50%
EEM -  0%
VMQ -  0%

```{r}
initial_wealth = 100000
set.rseed(1)  #Setting seed
sim1 = foreach(i=1:5000, .combine='rbind') %do% {
  total_wealth = initial_wealth
  weights = c(.2, 0.3, 0.5, 0, 0)
  holdings = weights * total_wealth
  n_days = 20
  wealthtracker = rep(0, n_days)
  for(today in 1:n_days) {
    return.today = resample(all_returns, 1, orig.ids=FALSE)
    holdings = holdings + holdings*return.today
    total_wealth = sum(holdings)
    wealthtracker[today] = total_wealth
    weights = c(.2, 0.3, 0.5, 0, 0)
    holdings = weights * total_wealth
  }
  wealthtracker
}

head(sim1)

```


Profit/loss for this portfolio
```{r}
par(bg = "gray")
mean(sim1[,n_days])
hist(sim1[,n_days]- initial_wealth, breaks=30,main = "Histogram of Portfolio Returns",xlab = "End of Period Return",ylab = "Frequency",col = "blue")
abline(v = mean(sim1[,n_days]) - initial_wealth,col = "red",lw = 3)
```

Calculating 5% value at risk
```{r}
quantile(sim1[,n_days], 0.05) - initial_wealth
```

Hence for this portfolio the average return is $100531 and value at risk is $3066.526

4) Next we will create an aggressive portfolio which increases our return as well as our risk. Ater trying several combinations, below is the weight we considered for this portfolio:

TLT -  0%
LQD -  0%
SPY -  0%
EEM - 90%
VMQ - 10%

```{r}
initial_wealth = 100000
set.rseed(1)      #Setting seed
sim1 = foreach(i=1:5000, .combine='rbind') %do% {
  total_wealth = initial_wealth
  weights = c(0, 0, 0, .9, .1)
  holdings = weights * total_wealth
  n_days = 20
  wealthtracker = rep(0, n_days)
  for(today in 1:n_days) {
    return.today = resample(all_returns, 1, orig.ids=FALSE)
    holdings = holdings + holdings*return.today
    total_wealth = sum(holdings)
    wealthtracker[today] = total_wealth
    weights = c(0, 0, 0, .9, .1)
    holdings = weights * total_wealth
  }
  wealthtracker
}

head(sim1)

```

Profit/loss for this portfolio
```{r}
par(bg = "gray")
mean(sim1[,n_days])
hist(sim1[,n_days]- initial_wealth, breaks=30,main = "Histogram of Portfolio Returns",xlab = "End of Period Return",ylab = "Frequency",col = "blue")
abline(v = mean(sim1[,n_days]) - initial_wealth,col = "red",lw = 3)
```

Calculating 5% value at risk
```{r}
quantile(sim1[,n_days], 0.05) - initial_wealth
```
Hence for this portfolio the average return is $101872.8 and value at risk is $12925.12 

Below is the summary of our results:

Portfolio 1 (Even split):
Mean Return - $100946.6
Value at Risk - $6297.993

Portfolio 2 (Safe):
Mean Return - $100531
Value at Risk - $3066.526 

Portfolio 3 (Risky):
Mean Return $101872.8
Value at Risk - 12925.12 

