---
title: "Modling Homework One"
author: "Ali Prasla"
date: "August 9, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part A.
$RC = Random Choice Event$

$TC = True Choice Event$

$Y = "Yes" Choice Event$

$N = "No" Choice Event$

Solve for $P(Y/TC)$

$P(Y/TC) = P(Y n TC) / P(TC)$

$P(TC) = .7$

$P(Y n TC) = P(Y) - P(Y n RC)$

$P(Y) = .65$

$P(Y n RC) = P(RC) * P(Y/RC)$

$P(RC) = .3$

$P(Y/RC) = .5$

$P(Y n RC) = $
```{r}
YAndRC = .3 * .5
YAndRC
```

$P(Y n TC) = $
```{r}
YAndTC = .65 - YAndRC
YAndTC
```

Finally,

$P(Y/TC) =  $
```{r}
YAndTC/.7
```

## Part B.
$P$ = Positive Test Event
$D$ = Has Disease Event

Solve for $P(D/P)$

$P(D/P) = (P(P/D) * P(D))/P(P)$



$P(P/D) =$ .993
$P(D) = $.000025
$P(P) = P(P n D) + P(P n D^c)$

$P(P n D) = P(P/D) * P(D)$
```{r}
PAndD = .993 * .000025
```

$P(P n D^c) = P(P/D^c) * P(D^c)$
$P(P/D^c) = 1 - .9999$
```{r}
PAndNotD = (1-.9999) * (1-.000025)
```

```{r}
P = PAndD + PAndNotD
```

Print Final Probability:
```{r}
(.993 * .000025) / P
```

Yes. I foresee problems with universial testing. This test effectively has a 20% false positive rate. Yet, to determine whether or not to implement this policy, it is important to look at other factors, like the deadliness of the disease and the efficacy of early treatment. 